Accurate interactions of cations with neutral and negatively charged membranes via combination of experiments and molecular simulation

### 2018-05-14 Manuscript/manuscript.tex
Written a paragraph on population analysis using number of contacts, updated figures, now all simulations with Ca2+ are at least 500ns, added TODO latex package.

\usepackage{rotating}  
\usepackage{times} 
\usepackage{graphicx} 
\usepackage{setspace} 
\usepackage{amsmath} 
\usepackage{epstopdf} 
\usepackage{csquotes} 
\usepackage{mhchem} 
\usepackage{chemfig}

@@

Where do the cations bind? Preferentially to POPS. Discuss further \dots

\begin{table}[tb!] 
  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$), 
    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$) 
    in different mixed POPC:POPS (5:1) bilayer models. 
  \label{tab:binding}} 
  \begin{tabular}{l|c c | c | c c} 
    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\ 
    \hline 
    ECC-POPS               &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.01 $   &  ??\%  &    ??\%    \\ 
    Lipid14/Dang           &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.03 $   &  ??\%  &    ??\%    \\ 
  \end{tabular} 
\end{table}

>>>>

\usepackage{rotating}  
\usepackage{times} 
\usepackage{graphicx} 
\usepackage{setspace} 
\usepackage{amsmath} 
\usepackage{epstopdf} 
\usepackage{csquotes} 
\usepackage{mhchem} 
\usepackage{chemfig} 
\usepackage[obeyFinal]{easy-todo}

@@

\textbf{Where do the cations bind?} 
The affinity towards PC and PS lipids and some of their groups was evaluated 
by counting contacts between the cation and the oxygen atoms of the lipids
similarly as was done in \cite{melcr18}. 
The threshold for counting a contact was set to be $0.3 \mathrm{nm}$, 
which encompasses the first peak of radial distribution function between the cations and the oxygen atoms of the lipids. 
In Table~\ref{tab:binding}, there are percentages of the membrane-bound calcium population for various membrane moieties. 
Although the lipid ratio in the membrane is 5~PC:1~PS,
half of the total population of calcium cations is bound to PS lipids
with 7\% bound only to them. 
This corroborates the intrinsincally higher affinity of PS lipids to calcium cations compared to PC lipids. 
The population analysis also suggests 
that the calcium cations prefer to reside in the phosphate region of the membrane. 
Hence, the carboxylate group in PS is coordinated to the cations bound in that region. 
The increased response of head group order parameters $\alpha$ and $\beta$ of PC 
in mixed 5~PC:1~PS bilayer compared to pure PC
is due to higer affinity of the membrane mediated by even a relatively small fraction (1/6) of PS lipids. 
This is in line with the steep onset of the response of the PS head group order parameters at lower concentrations.
The complex response of the head group order parameters of PS lipids 
is due to the confomational change of the carboxylate group that is attracted more towards the phosphate region. 
In order to examine whether the carboxylate and phosphate groups in PS are bound 
to the same calcium cation or rather different cations, 
more detailed analysis is required. 
\todo{Detailed analysis of the interactions of cations and PC/PS head groups using MSM.}

\begin{table}[tb!] 
  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), 
           and the percentages of Ca$^{2+}$ bound to 
           either to PC or PS, 
           to PC or PS without phosphate group,
           to PC or PS without carboxylate group,
           to PC and PS at the same time,
           only to PC, 
           and only to PS 
           in a mixed POPC:POPS (5:1) bilayer. 
          \label{tab:binding}} 
  \begin{tabular}{l| c | c c c c c c } 
    model                  & $C_{ion}\,/\,\mathrm{mM}$ & PC+PS  &  PC+PS no PO$_4$ & PC+PS no COO$^-$ & PC-$Ca^{2+}$-PS  & PC & PS  \\ 
    \hline 
    ECC-lipids             &  $240\pm 10 $             & 100    &   98-99          & 100              & 43               & 50 &  7  \\
    \hline 
    Lipid14/17             &  $280\pm 10 $             & 100    &   ---            & ---              & ---              & ---& --- \\
    \hline 
  \end{tabular} 
\end{table}
